### 2016-08-23 syntax.tex
syntax

\heading{Preliminaries. } When~$\univ$ is a set and $n>0$ is an
integer, we let $[\univ]^n$ denote the set of all size-$n$ subsets
%\jnote{Multisets?} 
of~$\univ$. We write $x \getsr \univ$ to denote
sampling an element from~$\univ$ and assigning this to~$x$, and we
extend this notation to randomized algorithms.  When~$X$ is a string
over some alphabet~$\Sigma$, we write $|X|$ for the bitlength
of~$X$, relative to some fixed encoding.  When~$X$ is a multiset, we
overload the notation so that~$|X|$ is the number of elements in~$X$
(with mutiplicity).

@@

\heading{Data structures.}  We being by defining a general primitive
that produces a representation of a given multiset, and provides a
mechanism to answer a specified collection of questions about that
multiset.  Informally, the representation should serve as a good
approximation of the multiset, at least with respect to the allowed
questions.
\begin{definition}[Data structure] \rm
Fix a non-empty set~$\univ$. A \emph{static data structure} (over universe~$\univ$)
is a tuple $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$ where:
\begin{itemize}
\item $\Sigma$ is the \emph{representation alphabet}. \jnote{Is
    there any reason not to simply fix this to $\{0,1\}$?}
\item $\Gamma$ is the set of possible \emph{query answers}.
    \jnote{Do we need this for any reason?}
\item $\calS$ is a set  of multisets over universe~$\univ$.
\item $\mathcal{Q}=\{q \colon \mathcal{S} \to \Gamma\}$ is a set
    of supported \emph{queries}.
\item $\Rep \colon \mathcal{S} \to \Sigma^* \times \bits^*
    \times \bits^*$ is a randomized \emph{representation
    algorithm}, taking as input $S \in \mathcal{S}$, and
    outputting a \emph{representation} $M \in \Sigma^*$,
    \emph{private side information} $\privaux \in \bits^*$, and
    \emph{public side information} $\pubaux \in \bits^*$.
\item $\Qry\colon \Sigma^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $M, \pubaux,\privaux, q$, and returns
    \emph{answer} $a \in \Gamma$. 
\end{itemize}
\hfill\dqed
\end{definition}
%\tsnote{I implemented this change.}\tsnote{I'm now leaning towards
%having $\Rep$ output~$M$ and explicit pubic/private auxillary data.
%Pushing secrets into~$M$ makes it more difficult to capture (for
%example) the Neidermeier et al.\ attacks in our privacy notions; see
%Figure~\ref{fig:privacy}.  In those attacks, the adversary would get
%the bitarray (which one naturally thinks of as~$M$) but not the
%secret keys for the PRFs/hash functions.  You could say that the
%bitarray is put in $\aux$ and that~$M$ \emph{only} contains the
%secret keys, but this is unnatural.} 
All of the data structures that
we will consider have the property that~$|M|$ does not depend on the
coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$. In
general this need not be true, but the assumption does not restrict
our syntax from capturing natural constructions (e.g., those used in
practice), and it makes it easier to state our security notions.

@@

We will sometimes say that a data structure is
$(1-\epsilon)$-correct, meaning it has error probability~$\epsilon$.
Note that $1-\epsilon$ is a worst-case lower bound on the
correctness of the data structure.

@@

\def\bin{{\sf Bin}}
\heading{``Trivial'' attacks, and error probability vs.\ error rate. }
Let~$\Pi$ be a scheme that is $(1-\epsilon)$-correct, and consider
an attack that asks random queries $q_1,q_2,\ldots,q_T$, i.e.,
uniform samples from~$\calQ$.
%The expected number of random queries needed to
%find a error is $1/\epsilon$ tries; thus, an attacker making $O(t/\epsilon)$ queries succeeds with constant probability (almost always).
%Actually, it seems worthwhile to formalize this.
Let $\mathrm{Err}_i$ be a random variable indicating the event $\Qry(M,\pubaux,\privaux,q_i)\neq q_i(S)$.
The $(1-\epsilon)$-correctness of~$\Pi$ guarantees that the probability that $\mathrm{Err}_i=1$ is at most~$\epsilon$,
and it is tempting to model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials with success probability~$\epsilon$.
However, a moment's reflection shows that this is not the case.  Concretely, consider the following scheme.  Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$.  For the queries, let $\calQ = \{q_x\colon \calS \to \bits \,|\, \forall x \in \univ\}$ where $\forall S \in \calS$ the predicate $q_x(S)=1 \Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random element~$s \in S$, and return $M = S \setminus \{s\}$ and $\privaux = \emptystring$.  To respond to queries, let~$\Qry(M,\pubaux,\privaux,q_x)=1$ iff $x \in M$.  It is easy to see that this scheme is $1/n$-correct; for any fixed~$S$, there are exactly~$n$ queries~$q_x$ that have probability~$1/n$  (over the coins of~$\Rep$) of causing a correctness error, and the remaining $|\univ|-n$ queries have probability zero of causing an error.  Moreover, once the coins of~$\Rep$ are fixed, there is exactly \emph{one} query that will cause an error.  Thus for random queries from~$Q$, for all~$i\in[T]$ we have $\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.

@@

\heading{Set-multiplicity Data Structures. } Our particular focus
will be on data structures that support multiplicity queries, of
which set-membership queries are a special case.
\begin{definition} A \emph{set-multiplicity data structure} is a data
structure with $\Gamma=\mathbb{N}$ and
where $\calQ=\{q_x \,|\, x \in \univ\}$ where $q_x(S)$ is defined to
be the multiplicity of~$x$ in the multiset~$S$. \hfill\dqed
\end{definition}
%
%\todo{Readdress issue of representation size.  A general statement might be something like `` The \emph{size} of a set-representation is the maximum (bit)length of any~$M$ that can be output by $\Rep$.''  But all of the natural constructions will have representation whose size will not depend on the coins of $\Rep$, i.e. they depend on $|S|$, number of hash functions, desired error rate. }

%Our treatment of $M,\privaux$ is not symmetric, since we only count $M$ against the size
%of the set-representation. When we define security
%we may give the attacker $\privaux$ but not~$M$, and in that case the set-representation cannot
%trivially make everything part of~$\privaux$.
%\tsnote{Encoding secrets into~$M$ may cause some headaches in the security definitions; we'll see. }
%\tsnote{By the way, I'm coming around to the
%  viewpoint that it is fair to include in the ``size'' of the
%  representation any secret information.  In makes more difficult
%  comparisons to classical lower-bounds (which [NY] makes, incorrectly
% IMO), but it does seem fair to say ``this is how many bits of memory
% you need to allocate to represent this set \emph{securely}.''}

%\heading{Bloom Filters and Related Data Structures. }
It is not hard to see Bloom filters are one possible instantiation a
set-multiplicity data structure.  Fix $\Sigma=\Gamma=\bits$.  Define
algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a
description of a hash function family~$\mathcal{H}=\{h \colon \univ
\to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions
$h_1, \ldots, h_k$ from~$\mathcal{H}$ and then compute an $m$-bit
array~$M$ (intialized to 0 everywhere) by setting $M[h_i(x)]=1$ for
all $i\in [k]$ and $x \in S$.  Finally, $\Rep$ returns
$\privaux=\emptystring$, $\pubaux = (h_1,\ldots, h_k)$, and~$M$ as the
representation.  
%
The algorithm~$\Qry$ is defined so that, on input $(M,\pubaux,\privaux,q_x)$, it
returns the minimum of the values $M[h_1(x)],\ldots,M[h_k(x)]$.  We note that, while our error
probability and error rate do not distinguish among ``types'' of
errors, they also do not preclude instantiations (like the classical
Bloom filter) that admit only one type of error, e.g.
false-positives.

>>>>

\heading{Preliminaries. } When~$\univ$ is a set and $n>0$ is an
integer, we let $[\univ]^n$ denote the set of all size-$n$ subsets
%\jnote{Multisets?} 
of~$\univ$, and $[[\univ]]^n$ for the set of all multisets
over~$\univ$ with total multiplicity~$n$. We write $x \getsr \univ$ to denote
sampling an element from~$\univ$ and assigning this to~$x$, and we
extend this notation to randomized algorithms.  When~$X$ is a string
over some alphabet~$\Sigma$, we write $|X|$ for the bitlength
of~$X$, relative to some fixed encoding.  When~$X$ is a multiset, we
overload the notation so that~$|X|$ is the total multiplicity of~$X$.

@@

\heading{Data structures.}  We being by defining a general primitive
that produces a representation of a given multiset, and provides a
mechanism to answer a specified collection of questions about that
multiset.  Informally, the representation should serve as a good
approximation of the multiset, at least with respect to the allowed
questions.
\begin{definition}[Data structure] \rm
Fix a non-empty set~$\univ$. A \emph{data structure} (over universe~$\univ$)
is a tuple $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$ where:
\begin{itemize}
\item $\Sigma$ is the \emph{representation alphabet}.
%\jnote{Is there any reason not to simply fix this to $\{0,1\}$?}
\item $\Gamma$ is the set of possible \emph{query answers}.
%\jnote{Do we need this for any reason?}
\item $\calS$ is a set  of multisets over universe~$\univ$.
\item $\mathcal{Q}=\{q \colon \mathcal{S} \to \Gamma\}$ is a set
    of supported \emph{queries}.
\item $\Rep \colon \mathcal{S} \to \Sigma^* \times \bits^*
    \times \bits^*$ is a randomized \emph{representation
    algorithm}, taking as input $S \in \mathcal{S}$, and
    outputting a \emph{representation} $M \in \Sigma^*$,
    \emph{private side information} $\privaux \in \bits^*$, and
    \emph{public side information} $\pubaux \in \bits^*$.  We write
    $(M,\pubaux,\privaux \getsr \Rep(S)$ as shorthand for this execution.
\item $\Qry\colon \Sigma^* \times \bits^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $M\in\Sigma^*$, $\pubaux,\privaux \in
    \bits^*$ and $q \in \calQ$, and outputting
    an \emph{answer} $a \in \Gamma$.  We write $a \gets
    \Qry(M,\pubaux,\privaux,q)$ for this execution.
\end{itemize}
\hfill\dqed
\end{definition}
%\tsnote{I implemented this change.}\tsnote{I'm now leaning towards
%having $\Rep$ output~$M$ and explicit pubic/private auxillary data.
%Pushing secrets into~$M$ makes it more difficult to capture (for
%example) the Neidermeier et al.\ attacks in our privacy notions; see
%Figure~\ref{fig:privacy}.  In those attacks, the adversary would get
%the bitarray (which one naturally thinks of as~$M$) but not the
%secret keys for the PRFs/hash functions.  You could say that the
%bitarray is put in $\aux$ and that~$M$ \emph{only} contains the
%secret keys, but this is unnatural.} 
All of the data structures that
we will consider have the property that~$|M|$ does not depend on the
coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$. In
general this need not be true, but the assumption does not restrict
our syntax from capturing natural constructions (e.g., those used in
practice), and it makes it easier to state our security notions.

@@

We will sometimes say that a data structure is
$(1-\epsilon)$-correct, meaning it has error probability~$\epsilon$.
Note that $1-\epsilon$ is a worst-case lower bound on the
correctness of the data structure. \tsnote{Don't know if this will be necessary.}

@@

\def\bin{{\sf Bin}}
\heading{``Trivial'' attacks, and error probability vs.\ error rate. }
Let~$\Pi$ be a scheme that is $(1-\epsilon)$-correct, and consider
an attack that asks random queries $q_1,q_2,\ldots,q_T$, i.e.,
uniform samples from~$\calQ$.
Let $\mathrm{Err}_i$ be a random variable indicating the event $\Qry(M,\pubaux,\privaux,q_i)\neq q_i(S)$.
The $(1-\epsilon)$-correctness of~$\Pi$ guarantees that the probability that $\mathrm{Err}_i=1$ is at most~$\epsilon$,
and it is tempting to model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials with success probability~$\epsilon$.
However, a moment's reflection shows that this is not the case.  Concretely, consider the following scheme.  Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$.  For the queries, let $\calQ = \{q_x\colon \calS \to \bits \,|\, \forall x \in \univ\}$ where $\forall S \in \calS$ the predicate $q_x(S)=1 \Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random element~$s \in S$, and return $M = S \setminus \{s\}$ and $\privaux = \emptystring$.  To respond to queries, let~$\Qry(M,\pubaux,\privaux,q_x)=1$ iff $x \in M$.  It is easy to see that this scheme is $1/n$-correct; for any fixed~$S$, there are exactly~$n$ queries~$q_x$ that have probability~$1/n$  (over the coins of~$\Rep$) of causing a correctness error, and the remaining $|\univ|-n$ queries have probability zero of causing an error.  Moreover, once the coins of~$\Rep$ are fixed, there is exactly \emph{one} query that will cause an error.  Thus for random queries from~$Q$, for all~$i\in[T]$ we have $\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.

@@

\heading{Set-multiplicity Data Structures. } Our particular focus
will be on data structures that support multiplicity queries, of
which set-membership queries are a special case.
\begin{definition} A \emph{set-multiplicity data structure} is a data
structure with $\Gamma=\mathbb{N}$ and
where $\calQ=\{q_x \,|\, x \in \univ\}$ where $q_x(S)$ is defined to
be the multiplicity of~$x$ in the multiset~$S$. \hfill\dqed
\end{definition}

It is not hard to see Bloom filters are an instantiation a
set-multiplicity data structure.  Fix $\Sigma=\Gamma=\bits$.  Define
algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a
description of a hash function family~$\mathcal{H}=\{h \colon \univ
\to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions
$h_1, \ldots, h_k$ from~$\mathcal{H}$ and then compute an $m$-bit
array~$M$ (intialized to 0 everywhere) by setting $M[h_i(x)]=1$ for
all $i\in [k]$ and $x \in S$.  Finally, $\Rep$ returns
$\privaux=\emptystring$, $\pubaux = (h_1,\ldots, h_k)$, and~$M$ as the
representation.  
%
The algorithm~$\Qry$ is defined so that, on input $(M,\pubaux,\privaux,q_x)$, it
returns the minimum of the values $M[h_1(x)],\ldots,M[h_k(x)]$.  We note that, while our error
probability and error rate do not distinguish among ``types'' of
errors, they also do not preclude instantiations (like the classical
Bloom filter) that admit only one type of error, e.g.
false-positives.
